Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2014-04-11 parts/ccl.tex
fixes after Morgan's review

We also demonstrated that general Process Hitting models with arbitrary priority
classes can also be encoded in the particular class of two-prioritized Process Hitting.
Therefore, our static analysis is applicable to a broad range of automata
networks with priorities.

>>>>

We also demonstrated that general Process Hitting models with arbitrary priority
classes can also be encoded in the particular class of two-prioritised Process Hitting.
Therefore, our static analysis is applicable to a broad range of automata
networks with priorities.

### 2014-04-11 parts/def-ph.tex
fixes after Morgan's review

\begin{example}
  \pref{fig:ph-livelock} gives an example of PH with $2$ classes of priorities where:
  \begin{align*}
    \PHs &= \{ a, b, c, ab \} \enspace, \\
    \PHl_a &= \{ a_0, a_1 \} \enspace, & \PHl_b &= \{ b_0, b_1 \} \enspace, \\
    \PHl_c &= \{ c_0, c_1 \} \enspace, & \PHl_{ab} &= \{ ab_{00}, ab_{01}, ab_{10}, ab_{11} \} \enspace.
  \end{align*}
  There also is especially: $\{ \PHhit{ab_{11}}{c_0}{c_1}, \PHhit{a_1}{a_1}{a_0}, \PHhit{a_0}{b_0}{b_1} \} \subseteq \PHh^{(2)}$.

@@

The construction of cooperation in PH allows to encode any Boolean function between cooperating processes \cite{PMR10-TCSB}.
Due to the introduction of priorities into the PH framework,
it is possible to build cooperations with no temporal shift by defining actions updating the cooperative sorts with the highest class of priority.
This allows to gain the same expressivity in PH than in Boolean networks, as stated in \pref{sec:dn}.
The aim of this paper is to allow the static analysis of the dynamics to be handled on PH models comprising such prioritised actions updating cooperative sorts.

@@

In the scope of this paper, we focus on a specific class of PH models
in which actions of $\PHh^{(1)}$ are called \emph{prioritized actions} and allow to model non-biological actions.
We consider in this section a PH model with $k$ classes of priorities: $\PH = (\PHs; \PHl; \PHa^{\langle k \rangle})$, with $k \in \sNN$
and we define here the restrictions that lead to this class of models.

@@

We note that a sort $a$ that is hit by no action (\ie $\forall h \in \PHh, \PHsort(\target{h}) \neq a$)
does comply with both definitions of well-formed components and well-formed cooperative sorts.
In this case, such a sort can be arbitrarily affected to either of sets.
Such case may occur for “input” sorts (which state is initially fixed and do not evolve)
or some special cases of cooperative sorts produces by the translations of \pref{sec:encodings}.

>>>>

\begin{example}
  \pref{fig:ph-livelock} gives an example of PH with $2$ classes of priorities where:
  \begin{align*}
    \PHs &= \{ a, b, c, ab \} \enspace, \\
    \PHl_a &= \{ a_0, a_1 \} \enspace, & \PHl_b &= \{ b_0, b_1 \} \enspace, \\
    \PHl_c &= \{ c_0, c_1 \} \enspace, & \PHl_{ab} &= \{ ab_{00}, ab_{01}, ab_{10}, ab_{11} \} \enspace.
  \end{align*}
  Note that: $\{ \PHhit{ab_{11}}{c_0}{c_1}, \PHhit{a_1}{a_1}{a_0}, \PHhit{a_0}{b_0}{b_1} \} \subseteq \PHh^{(2)}$.

@@

The construction of cooperation in PH allows to encode any Boolean function between cooperating processes \cite{PMR10-TCSB}.
Due to the introduction of priorities into the PH framework,
it is possible to build cooperations with no temporal shift by defining actions updating the cooperative sorts with the highest class of priority.
This allows to gain the same expressivity in PH than in Boolean networks, as stated in \pref{sec:dn}.
The aim of this paper is to allow the static analysis of the dynamics to be handled on PH models involving such prioritised actions updating cooperative sorts.

@@

In the scope of this paper, we focus on a specific class of PH models
in which actions of $\PHh^{(1)}$ are called \emph{prioritised actions} and allow to model non-biological actions.
We consider in this section a PH model with $k$ classes of priorities: $\PH = (\PHs; \PHl; \PHa^{\langle k \rangle})$, with $k \in \sNN$
and we define here the restrictions that lead to this class of models.

@@

We note that a sort $a$ that is hit by no action (\ie $\forall h \in \PHh, \PHsort(\target{h}) \neq a$)
does comply with both definitions of well-formed components and well-formed cooperative sorts.
In this case, such a sort can be arbitrarily affected to either set.
Such a case may occur for “input” sorts (whose state is initially fixed and does not evolve)
or some special cases of cooperative sorts produced by the translations of \pref{sec:encodings}.

### 2014-04-11 parts/intro.tex
fixes after Morgan's review

Of course, such formal models on which analyses are performed are abstract representations of the actual studied systems.
They are associated with parameters that have to be synthesised %so as to be as much as possible fitting with the real systems having some observed behaviors.
to give the most faithful representation of the real systems with their observed behaviours.
As a matter of fact, the abstractions we get are more or less rough or accurate.
Usual formal frameworks for such modelling activities are state-transition systems or process algebras. % Petri nets
We developed a quite similar framework named the Process Hitting \cite{PMR10-TCSB},
consisting in a restriction where the evolution of a component is determined by the state of at most one other component that does not evolve.
In a sense, these kind of actions are of the form $X + Y \rightarrow X + Y'$ where $X$ behaves like a catalyst molecule that “hits” another molecule $Y$ and changes it into $Y'$, without being itself changed.
Assuming catalysts are always available, this can represent any biochemical system made of monomolecular reactions, and can also represent catalytic networks such as metabolic networks.
%, with the aim of avoiding to build the whole state space in order to be able to tackle very large systems
Our motivation behind this framework was to design a model and analysis techniques adapted to biological modelling.
These analyses avoid to build the whole state space, which allows to tackle very large systems (that would have led to a huge number of states, hopelessly too huge to be analysed).
They are based on the fact that most biological models have few levels of expression per component:
in Boolean networks~\cite{kauffman69,Thomas73} there are only two levels per component and in their multivalued equivalent, Asynchronous Discrete Networks~\cite{deJong02}, components rarely have more that four levels.

@@

This paper is an extension of \cite{FPMR13-CS2Bio} were we
introduced a new extension to the semantics of Process Hitting by partitioning actions into classes of priorities
(\pref{sec:ph}).
One of the objectives is to reach an accurate representation of cooperating components in the model, that was not fulfilled with the initial semantics.
In \pref{sec:sa},
we develop an efficient under-approximation of the reachability of the state of components on a subclass of this new framework, thus allowing to compute efficient static analysis.
This local reachability under-approximation can be also easily extended to study the reachability of a global state.
Furthermore, we provide a new extended under-approximation method to decide a successive reachability property.

@@

>>>>

Of course, such formal models on which analyses are performed are abstract representations of the actual studied systems.
They are associated with parameters that have to be synthesised %so as to be as much as possible fitting with the real systems having some observed behaviors.
to give the most faithful representation of the real systems with their observed behaviours.
As a matter of fact, the abstractions we get are more or less rough or accurate.
Usual formal frameworks for such modelling activities are state-transition systems or process algebras. % Petri nets
We developed a quite similar framework named the Process Hitting \cite{PMR10-TCSB},
consisting in a restriction where the evolution of a component is determined by the state of at most one other component that does not evolve.
In a sense, these kind of actions are of the form $X + Y \rightarrow X + Y'$ where $X$ behaves like a catalyst molecule that “hits” another molecule $Y$ and changes it into $Y'$, without being changed itself.
Assuming catalysts are always available, this can represent any biochemical system made of monomolecular reactions, and can also represent catalytic networks such as metabolic networks.
%, with the aim of avoiding to build the whole state space in order to be able to tackle very large systems
Our motivation behind this framework was to design a model and analysis techniques adapted to biological modelling.
These analyses avoid to build the whole state space, which allows to tackle very large systems (that would have led to a huge number of states, hopelessly too huge to be analysed).
They are based on the fact that most biological models have few levels of expression per component:
in Boolean networks~\cite{kauffman69,Thomas73} there are only two levels per component and in their multivalued equivalent, Asynchronous Discrete Networks~\cite{deJong02}, components rarely have more that four levels.

@@

This paper is an extension of \cite{FPMR13-CS2Bio} where we
introduced a new extension to the semantics of Process Hitting by partitioning actions into classes of priorities
(\pref{sec:ph}).
One of the objectives is to reach an accurate representation of cooperating components in the model, that was not fulfilled with the initial semantics, using only two classes of priorities.
In \pref{sec:sa},
we develop an efficient under-approximation of the reachability of the state of components on a subclass of this new framework, thus allowing to compute efficient static analysis.
This local reachability under-approximation can be also easily extended to study the reachability of a global state.
Furthermore, we provide a new extended under-approximation method to decide a successive reachability property.

@@

The main additions in this paper compared to \cite{FPMR13-CS2Bio} are
two theorems, the first
stating that any Process Hitting model with any number of classes of priorities can be represented
with only two classes of priorities,
and the other giving a method to cope with successive reachability,
both extending the application range of the static analysis we developed.

### 2014-04-11 parts/models-ph-php.tex
fixes after Morgan's review

\begin{definition}[Playability property language ($\F$)]
  \label{def:pp}
  A \emph{playability property} is an element of the language $\F$ inductively defined by:
  \begin{itemize}
    \item $\top$ and $\bot$ belong to $\F$;
    \item if $a \in \ov{\PHs}$ and $a_i \in \ov{\PHl}_a$, then $a_i \in \F$ and we call it an \emph{atom};
    \item if $P \in \F$ and $Q \in \F$, then $\neg P \in \F$, $P \wedge Q \in \F$ and $P \vee Q \in \F$.
  \end{itemize}
  If $P \in \F$ is a playability property and $\mysigma \in \PHsubl$ is a sub-state of $\oPH$,
  we note $\Fsem{P}{\mysigma}$ the \emph{evaluation} of $P$ in $\mysigma$:
  \begin{itemize}
    \item if $P = a_i \in \ov{\PHs}_a$ is an atom, with $a \in \ov{\PHs}$, then $\Fsem{a_i}{\mysigma}$ is true iff $a_i \in \mysigma$;
    \item if $P$ is not an atom, then $\Fsem{P}{\mysigma}$ is true iff all its subformulas are inductively true in $\mysigma$
      with the classical semantics for the logic operators $\top$, $\bot$, $\neg$, $\wedge$ and $\vee$.
  \end{itemize}
\end{definition}

Because we only use classical logic operators, the formulas of Boolean logic on 
distributivity, associativity and commutativity can be used, together with De Morgan's laws on negation.
We also have the following property for the negation of an atom:
\[\forall a \in \ov{\PHs}, \forall a_i \in \ov{\PHl}_a, \forall \mysigma \in \PHsubl,
  \Fsem{\neg a_i}{\mysigma} \Leftrightarrow \Fsem{\bigvee_{\substack{a_j \in \PHl_a\\a_j \neq a_i}} a_j}{\mysigma}\]
Indeed, if a process is not active in a state, this means that another process of the same sort is active.

In \pref{def:fop}, we define the the operator $\Fopsymbol$ which characterises the playability of an action
given the semantics of PH (\pref{def:play}).
This operator considers the \emph{virtual hitters} of an action,
which are the sub-states given by $\csState$ to activate the hitter of the given action if it is a process of cooperative sort.
Furthermore, the target of the action if not taken into account as it is not needed for the rest of the flattening.
%\pref{th:ppplay} gives the condition for an action to be playable given its playability property.
%
\begin{definition}[Playability property of an action ($\Fopsymbol : \PHh \rightarrow \F$)]\label{def:fop}
  \[\forall h \in \ov{\PHh}, \Fop{h} \equiv \virtualhitters(h) \wedge
    \left( \bigwedge_{\substack{g \in \PHh^{(n)}\\n < \prio(h)}}
    \neg \left( \target{g} \wedge \hitter{g}\right) \right)\]
  where if we note $\hitter{h} = a_i$, then:
  \[\virtualhitters(h) =
    \begin{cases}
      a_i & \text{if } a \in \components \text{ or } h \in \ov{\PHh}^{(1)} \\
      \bigvee_{ps \in \csState(a_i)} \left( \bigwedge_{p \in ps} p \right) & \text{if } a \in \cs
    \end{cases}\]
\end{definition}
%
By construction of this operator and given the dynamics of a PH model,
an action $h$ is playable in a state $s \in \ov{\PHl}$ if and only if: $\Fsem{\Fop{h}}{s} \wedge \target{h} \in s$.
%%
%\begin{theorem}
%\label{th:ppplay}
%  $h \in \ov{\PHh}$ is playable in $s \in \ov{\PHl}$ if and only if $\Fsem{\Fop{h}}{s} \wedge \target{h} \in s$.
%\end{theorem}
%%
%\newproof{proofppplay}{Proof of \pref{th:ppplay}}
%\begin{proofppplay}
%  By definition of the semantics of a PH given in \pref{def:play}.
%\end{proofppplay}

@@

With \pref{lem:ppplaysubset}, we can then characterise the playability of an action in a state only with a sub-state.
This sub-state corresponds to one of the conjunctions of its playability property's DNF.
Finally, \pref{def:flattening} gives the construction of the flattening of $\oPH$:
for each action $h \in \ov{\PHh}$, several cooperative sorts $f^{h,i}$ are built to reflect each of the conjunction in $\Fop{h}$,
\ie for $i \in \segm{1}{\n}$.
This construction allows to obtain the same dynamics as $\oPH$, as stated by \pref{th:bisimPHP}.
A detailed proof of this theorem is available in \pref{suppl:demoflattening}.
%
\begin{lemma}
\label{lem:ppplaysubset}
  Let $h \in \ov{\PHh}$ and $s \in \ov{\PHl}$;
  $h$ is playable in $s$ if and only if:
  \[\target{h} \in s \wedge \exists \mysigma \subseteq s, \Fsem{\Fop{h}}{\mysigma} \enspace.\]
%  \[\exists i \in \segm{1}{\n}, \exists \mysigma \in \bigtimes{b \in \PHdep{i}{h}} \PHl_b,
%    \mysigma \subseteq s \wedge \Fsem{\F(h)}{\mysigma} \enspace.\]
\end{lemma}
%
%\newproof{proofppplaysubset}{Proof of \pref{lem:ppplaysubset}}
%\begin{proofppplaysubset}
\begin{proof}
  ($\Rightarrow$)
    If $h$ is playable in $s$, then $\target{h} \in s$ and $\Fsem{\Fop{h}}{s}$.
    Thus, $\Fop{h} \not\equiv \bot$ and, by property of a DNF,
    at least one of the $\n$ conjunctions of $\Fop{h}$ is true in $s$.
    Suppose the $i$\textsuperscript{th} conjunction is true in $s$, with $i \in \segm{1}{\n}$;
    then we have: $\forall j \in \segm{1}{\m}, p_{i,j} \in s$.
%    Let $\mysigma \in \bigtimes{b \in \PHdep{i}{h}} \PHl_b$
    Let $\mysigma \in \PHsubl_{\PHdep{i}{h}}$
    with $\forall b \in \PHdep{i}{h}, \PHget{\mysigma}{b} = \PHget{s}{b}$.
    We immediately have: $\mysigma \subseteq s$,
    and, by construction of $\PHdep{i}{h}$, $\Fsem{\Fop{h}}{\mysigma}$.

  ($\Leftarrow$)
    $\Fsem{\Fop{h}}{\mysigma}$ therefore $\Fsem{\Fop{h}}{s}$; as $\target{h} \in s$, $h$ is playable in $s$.
\end{proof}
%\end{proofppplaysubset}

\begin{definition}
  \label{def:flattening}
  If $k \in \sNN$ and $\oPH = (\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}})$
  is a PH with $k$ classes of priorities that respects all Conditions of \pref{ssec:hypothesis},
  we note $\PHflat(\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}}) = (\PHs; \PHl; \PHa^{\angles{2}})$
  the \emph{flattening} of $\oPH$, where:
  \begin{itemize}
    \item $\PHs = \ov{\PHs} \cup \PHs_f$
      where $\PHs_f = \{ f^{h,i} \mid h \in \ov{\PHh} \wedge \n \geq 1 \wedge i \in \segm{1}{\n} \}$;
    \item $\PHl = \left( \bigtimes{a \in \ov{\PHs}} \PHl_{a} \right) \times
      \left(\bigtimes{f^{h,i} \in \PHs_f} \PHl_{f^{h,i}} \right)$,
      where $\forall a \in \ov{\PHs}, \PHl_{a} = \ov{\PHl}_{a}$, and
      $\forall f^{h,i} \in \PHs_f,
      \PHl_{f^{h,i}} = \{ f^{h,i}_\mysigma \mid \mysigma \in \PHsubl_{\PHdep{i}{h}} \}$;
    \item $\PHh^{(1)} = \{ \PHhit{a_k}{f^{h,i}_\mysigma}{f^{h,i}_{\mysigma'}} \mid
      h \in \ov{\PHh} \wedge
      %\n \geq 1 \wedge i \in \segm{1}{\n} \wedge
      f^{h,i} \in \PHs_f \wedge
      a \in \PHdep{i}{h} \wedge a_k \in \PHl_a \wedge
      f^{h,i}_\mysigma , f^{h,i}_{\mysigma'} \in \PHl_{f^{h,i}} \wedge
      \PHget{\mysigma}{a} \neq a_k \wedge \mysigma' = \mysigma \Cap \{ a_k \} \}$;
%      \todo{From here: ...} \PHget{\mysigma'}{a} = a_k \wedge
%      (\forall b \in \PHdep{i}{h} \setminus \{ a \}, \PHget{\mysigma'}{b} = \PHget{\mysigma}{b}) \}$;
%      \todo{... can be replaced by: $\mysigma' = \mysigma \Cap \{ a_k \}$}
    \item $\PHh^{(2)}=\{ \PHhit{f^{h,i}_\mysigma}{\target{h}}{\bounce{h}} \mid
      h \in \ov{\PHh} \setminus \ov{\PHh}^{(1)} \wedge
      f^{h,i} \in \PHs_f \wedge
      f^{h,i}_\mysigma \in \PHl_{f^{h,i}} \wedge \Fsem{\Fop{h}}{\mysigma} \}$.
  \end{itemize}
  Given a state $s \in \PHl$, $\unflats{s} = \os$ is the corresponding state in $\ov{\PHl}$:
  $\forall a \in \ov{\PHs}, \PHget{\os}{a} = \PHget{s}{a}$.

>>>>

\begin{definition}[Playability property language ($\F$)]
  \label{def:pp}
  A \emph{playability property} is an element of the language $\F$ inductively defined by:
  \begin{itemize}
    \item $\top$ and $\bot$ belong to $\F$;
    \item if $a \in \ov{\PHs}$ and $a_i \in \ov{\PHl}_a$, then $a_i \in \F$ and we call it an \emph{atom};
    \item if $P \in \F$ and $Q \in \F$, then $\neg P \in \F$, $P \wedge Q \in \F$ and $P \vee Q \in \F$.
  \end{itemize}
  If $P \in \F$ is a playability property and $\mysigma \in \PHsubl$ is a sub-state of $\oPH$,
  we note $\Fsem{P}{\mysigma}$ the \emph{evaluation} of $P$ in $\mysigma$:
  \begin{itemize}
    \item if $P = a_i \in \ov{\PHl}_a$ is an atom, with $a \in \ov{\PHs}$, then $\Fsem{a_i}{\mysigma}$ is true iff $a_i \in \mysigma$;
    \item if $P$ is not an atom, then $\Fsem{P}{\mysigma}$ is true iff all its subformulas are inductively true in $\mysigma$
      with the classical semantics for the logic operators $\top$, $\bot$, $\neg$, $\wedge$ and $\vee$.
  \end{itemize}
\end{definition}

Because we only use classical logic operators, the formulas of Boolean logic on 
distributivity, associativity and commutativity can be used, together with De Morgan's laws on negation.
We also have the following property for the negation of an atom:
\[\forall a \in \ov{\PHs}, \forall a_i \in \ov{\PHl}_a, \forall \mysigma \in \PHsubl,
  \Fsem{\neg a_i}{\mysigma} \Leftrightarrow \Fsem{\bigvee_{\substack{a_j \in \ov{\PHl}_a\\a_j \neq a_i}} a_j}{\mysigma}\]
Indeed, if a process is not active in a state, this means that another process of the same sort is active.

In \pref{def:fop}, we define the the operator $\Fopsymbol$ which characterises the playability of an action
given the semantics of PH (\pref{def:play}).
This operator considers the \emph{virtual hitters} of an action,
which are the sub-states given by $\csState$ to activate the hitter of the given action if it is a process of cooperative sort.
Furthermore, the target of the action is not taken into account as it is not needed for the rest of the flattening.
%
\begin{definition}[Playability property of an action ($\Fopsymbol : \PHh \rightarrow \F$)]\label{def:fop}
  \[\forall h \in \ov{\PHh}, \Fop{h} \equiv \virtualhitters(h) \wedge
    \left( \bigwedge_{\substack{g \in \ov{\PHh}^{(n)}\\n < \prio(h)}}
    \neg \left( \target{g} \wedge \hitter{g}\right) \right)\]
  where if we note $\hitter{h} = a_i$, then:
  \[\virtualhitters(h) =
    \begin{cases}
      a_i & \text{if } a \in \components \text{ or } h \in \ov{\PHh}^{(1)} \\
      \bigvee_{ps \in \csState(a_i)} \left( \bigwedge_{p \in ps} p \right) & \text{if } a \in \cs
    \end{cases}\]
\end{definition}
%
By construction of this operator and given the dynamics of a PH model,
an action $h$ is playable in a state $s \in \ov{\PHl}$ if and only if: $\Fsem{\Fop{h}}{s} \wedge \target{h} \in s$.

@@

With \pref{lem:ppplaysubset}, we can then characterise the playability of an action in a state only with a sub-state.
This sub-state corresponds to one of the conjunctions of its playability property's DNF.
Finally, \pref{def:flattening} gives the construction of the flattening of $\oPH$:
for each action $h \in \ov{\PHh}$, several cooperative sorts $f^{h,i}$ are built to reflect each of the conjunction in $\Fop{h}$,
\ie for $i \in \segm{1}{\n}$.
This construction allows to obtain the same dynamics as $\oPH$, as stated by \pref{th:bisimPHP}.
A detailed proof of this theorem is available in \pref{suppl:demoflattening}.
%
\begin{lemma}
\label{lem:ppplaysubset}
  Let $h \in \ov{\PHh}$ and $s \in \ov{\PHl}$;
  $h$ is playable in $s$ if and only if:
  \[\target{h} \in s \wedge \exists \mysigma \subseteq s, \Fsem{\Fop{h}}{\mysigma} \enspace.\]
\end{lemma}
%
\begin{proof}
  ($\Rightarrow$)
    If $h$ is playable in $s$, then $\target{h} \in s$ and $\Fsem{\Fop{h}}{s}$.
    Thus, $\Fop{h} \not\equiv \bot$ and, by property of a DNF,
    at least one of the $\n$ conjunctions of $\Fop{h}$ is true in $s$.
    Suppose the $i$\textsuperscript{th} conjunction is true in $s$, with $i \in \segm{1}{\n}$;
    then we have: $\forall j \in \segm{1}{\m}, p_{i,j} \in s$.
    Let $\mysigma \in \PHsubl_{\PHdep{i}{h}}$
    with $\forall b \in \PHdep{i}{h}, \PHget{\mysigma}{b} = \PHget{s}{b}$.
    We immediately have: $\mysigma \subseteq s$,
    and, by construction of $\PHdep{i}{h}$, $\Fsem{\Fop{h}}{\mysigma}$.

  ($\Leftarrow$)
    $\Fsem{\Fop{h}}{\mysigma}$ therefore $\Fsem{\Fop{h}}{s}$; as $\target{h} \in s$, $h$ is playable in $s$.
\end{proof}

\begin{definition}
  \label{def:flattening}
  If $k \in \sNN$ and $\oPH = (\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}})$
  is a PH with $k$ classes of priorities that respects all Conditions of \pref{ssec:hypothesis},
  we note $\PHflat(\ov{\PHs}; \ov{\PHl}; \ov{\PHa}^{\angles{k}}) = (\PHs; \PHl; \PHa^{\angles{2}})$
  the \emph{flattening} of $\oPH$, where:
  \begin{itemize}
    \item $\PHs = \ov{\PHs} \cup \PHs_f$
      where $\PHs_f = \{ f^{h,i} \mid h \in \ov{\PHh} \wedge \n \geq 1 \wedge i \in \segm{1}{\n} \}$;
    \item $\PHl = \left( \bigtimes{a \in \ov{\PHs}} \PHl_{a} \right) \times
      \left(\bigtimes{f^{h,i} \in \PHs_f} \PHl_{f^{h,i}} \right)$,
      where $\forall a \in \ov{\PHs}, \PHl_{a} = \ov{\PHl}_{a}$, and
      $\forall f^{h,i} \in \PHs_f,
      \PHl_{f^{h,i}} = \{ f^{h,i}_\mysigma \mid \mysigma \in \PHsubl_{\PHdep{i}{h}} \}$;
    \item $\PHh^{(1)} = \{ \PHhit{a_k}{f^{h,i}_\mysigma}{f^{h,i}_{\mysigma'}} \mid
      h \in \ov{\PHh} \wedge
      f^{h,i} \in \PHs_f \wedge
      a \in \PHdep{i}{h} \wedge a_k \in \PHl_a \wedge
      f^{h,i}_\mysigma , f^{h,i}_{\mysigma'} \in \PHl_{f^{h,i}} \wedge
      \PHget{\mysigma}{a} \neq a_k \wedge \mysigma' = \mysigma \Cap \{ a_k \} \}$;
    \item $\PHh^{(2)}=\{ \PHhit{f^{h,i}_\mysigma}{\target{h}}{\bounce{h}} \mid
      h \in \ov{\PHh} \setminus \ov{\PHh}^{(1)} \wedge
      f^{h,i} \in \PHs_f \wedge
      f^{h,i}_\mysigma \in \PHl_{f^{h,i}} \wedge \Fsem{\Fop{h}}{\mysigma} \}$.
  \end{itemize}
  Given a state $s \in \PHl$, $\unflats{s} = \os$ is the corresponding state in $\ov{\PHl}$:
  $\forall a \in \ov{\PHs}, \PHget{\os}{a} = \PHget{s}{a}$.
